Robust Subspace and Higher Moments Distances for Botnet Detection

### 2019-04-27 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
improvements on literature review

\begin{abstract}
Botnets... In this context, signal processing techniques have been applied to attack detection due to their capability to detect anomalies that are previously unknown, i.e. blind detection, in contaminated or imbalanced data.... This paper proposes a cumulant-based robust estimates for network attack detection, by combining kurtosis and skewess with Minimum Covariance Determinant. In order to validate the proposed framework, we consider CTU-13 dataset. The experiments show that the proposed method achieves satisfactory levels of precisio, recall and F1 score for network attack detection.
\end{abstract}

@@

Botnets have emerging as a key threat for cyber security, considering that Botnets are the key platform for many Internet attacks (i.e. spam, distributed denial-of-service (DDoS), identity theft, and phishing). Most of the current botnet detection approaches work only on specific botnet command and control (. Lots of efforts are directed to avoid attacks from known attackers, however attacks from one source is not real nowadays, since the attacks are usualy from distributed botnets that change their approaches frequently, in order to avoid behavioral or pattern based detecions. This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,da2018online,Wang2018ddosbotnets}.

@@

Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning}. Learning algorithms for imbalanced data has been a challenging research topic. The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}. He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning. Some techniques can make the data discriminative for better discriminative analytics: Data engineering (i.e. sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, acarali2016survey}); Feature extraction (i.e. decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}). To face a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Works}
\label{sec:relatedworks}

@@

Khattak \emph{et al} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, based on a decision module as a Complex Event Processing (CEP) engine and based on a correlation framework that continuously receives events from the sensor module and correlates them according to rule(s).

@@

\cite{Wang2018ddosbotnets} present an in-depth analysis based on 50,704 different Internet DDoS attacks launched by 674 botnets from 23 different botnet families. Geospatial distribution of the attacking sources follows certain patterns, multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.

\cite{wang2017botnet} (Botnet detection based on anomaly and community detection)

\cite{da2018online} (Online detection of Botnets on Network Flows using Stream Mining)

\cite{moustafa2019holistic} (A holistic review of Network Anomaly Detection Systems: A comprehensive survey). Network anomaly detection approaches are classified by \cite{moustafa2019holistic} in six categories: classification, clustering, deep learning, knowledge, combination and statistical-based.

@@

\cite{acarali2016survey} (Survey of approaches and features for the identification of HTTP-based botnet traffic)

@@

@@

To deal with the data imbalance of CTU-13 data set, we adopt a sampling method based on data aggragation \cite{acarali2016survey} for feature generation, in order to obtain less imbalanced and more discrimnative data.

Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15, 0.25, 1 and 2 seconds, for each scenario of the CTU-13 data set.

@@

		In anomalous detection problems, where anomalies are rare events, if we classify all observations as normal and apply an accuracy evaluation, we would have high accuracy but poor true-positive detection.
		Due to the importance given by F-measure to true-positive detection \cite{moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.

>>>>

\begin{abstract}
Botnets... In this context, signal processing techniques have been applied to attack detection due to their capability to detect anomalies that are previously unknown, i.e. blind detection, in contaminated or imbalanced data.... This paper proposes a cumulant-based robust estimates for network attack detection, by combining kurtosis and skewess with Minimum Covariance Determinant. In order to validate the proposed framework, we consider CTU-13 dataset. The experiments show that the proposed method achieves satisfactory levels of precisio, recall and F1 score for network attack detection. We apply our method to real-world botnet traffic and compare its performance with other methods.
\end{abstract}

@@

Botnets have emerging as a key threat for cyber security, considering that Botnets are the key platform for many Internet attacks (i.e. spam, distributed denial-of-service (DDoS), identity theft, and phishing). Most of the current botnet detection approaches work only on specific botnet command and control (. Lots of efforts are directed to avoid attacks from known attackers, however attacks from one source is not real nowadays, since the attacks are usualy from distributed botnets that change their approaches frequently, in order to avoid behavioral or pattern based detecions. This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,da2018online,Wang2018ddosbotnetssurvey}.

@@

Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning}. Learning algorithms for imbalanced data has been a challenging research topic. The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}. He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning. Some techniques can make the data discriminative for better discriminative analytics: Data engineering (i.e. sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, garcia2014empirical, acarali2016survey}); Feature extraction (i.e. decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}). To face a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% network anomaly detectiom, botnet detection, botnet detection approaches, approaches based on subspace learning, sampling and aggregation, about the methodos to compare and metrics
\section{Literature Review}
\label{sec:relatedworks}

Talk about network anomaly detection...

Bhuyan \emph{et al} \cite{acarali2016survey} provide an overview of facets of network anomaly detection, present attacks normally encountered by network intrusion detection systems, and categorize existing network anomaly detection methods and systems based on the underlying techniques. Ahmed Bhuyan \emph{et al} \cite{acarali2016survey} present an analysis of four major categories of anomaly detection techniques which include classification, statistical, information theory and clustering. Recently, Moustafa \emph{et al} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark datasets for training and validating approaches for detection.

Acarali \emph{et al} \cite{acarali2016survey} surveys recent network-based detection approaches for HTTP-based botnets, surveys traffic-based features used to detect bot traffic and presents an abstraction of the main types of features in relation to protocols and OSI layers. Wang \emph{et al} \cite{Wang2018ddosbotnetssurvey} present an analysis based on 50,704 different Internet DDoS attacks originated of 674 botnets from 23 different botnet families with a total of 9,026 victim belonging to 1,074 organizations in 186 countries. Their analysis reveals that geolocation of the attacking sources follows patterns and enables source prediction, and highlight that multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, also presents that there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.

Khattak \emph{et al} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, based on a Complex Event Processing (CEP) engine and on a correlation framework that continuously receives events and correlates them according to rule(s). BotFlex's results are compared to BotHunter, but the evaluation relies in a own and not public data set.

Wang and Paschalidis \cite{wang2017botnet} propose an botnet detection approach based on anomaly and community detection, aiming for detecting botnets and identifying bots before the botnet becomes active. The first stage detects anomalies by leveraging large deviations of an empirical distribution. The second stage detects the bots using ideas from social network community detection in a graph that captures correlations of interactions among nodes over time. This work is compared with the BotHunter \cite{gu2007bothunter} on the CTU-13 botnet dataset \cite{garcia2014empirical}.

\cite{da2018online} (Online detection of Botnets on Network Flows using Stream Mining) Ver se podemos calcular o F1 e comparar os resutados, mas não seria justo dado que a abordagem deles nao implementa o mesmo sampling e é supervisionado

@@

@@

@@

@@

\cite{garcia2014empirical} proposed CTU-13 dataset and argues that for evaluations of anomaly detection, none of the botnet families used in the training and crossvalidation dataset should be used in the testing dataset, considering that this ensures that the methods can generalize and detect new behaviors. However, hour proposal is semi-supervised and does not rely on training data with anomalies. Adopting the training and testing approach of \cite{garcia2014empirical}, some botnets malware wouldnt be evaluated, since in Garcia's proposal some botnets are present only for training. Therefore, we evaluate each scenario individually.

A. Description of CTU-13 Dataset
The CTU-13 is a dataset of botnet traffic that was captured
in the Czech Technical University [24]. It contains 13 scenarios
with various botnet types. We consider scenarios 1, 2, 6, 8, and
9 of the dataset.
• Scenario 1 corresponds to an IRC-based botnet that sent
spam for almost 6.5 hours.
• Scenario 2 (2.5-hours) is from the same botnet.
• Scenario 6 is from a botnet that scanned SMTP (Simple
Mail Transfer Protocol) servers for 2 hours and connected
to several RDP (Remote Desktop Protocol) services. Dif-
ferent with Scenarios 1 and 2, this botnet neither sent
spam nor did it attack. It’s C&C server used a proprietary
protocol.
• Scenario 8 is from a botnet that contacted a lot of Chinese
C&C hosts and received large amounts of encrypted data.
The botnet also cracked the passwords of machines during
the 19-hour attack.
• Scenario 9 corresponds to a case where 10 local hosts
were infected by a spamming botnet. More than 600 spams
were sent over 5 hours.
For all the scenarios, the authors of the CTU-13 dataset
convert the captured pcap files to NetFlows and release the
processed flows. The dataset contains ground-truth labels for
flows as follows: flows from or to the infected machines
are labeled as “botnet”; flows from or to noninfected ma-
chines are labeled as “normal”; all other flows are labeled as
“background.”

@@

To deal with the data imbalance of CTU-13 data set, we adopt a sampling method based on data aggragation for feature generation, in order to obtain less imbalanced and more discrimnative data. Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15 and 0.25 seconds, for each scenario of the CTU-13 data set.

@@

In anomalous detection problems, where anomalies are rare events, if we classify all observations as normal and apply an accuracy evaluation, we would have high accuracy but poor true-positive detection.

Due to the importance given by F-measure to true-positive detection \cite{moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.
